---
title: "Build a Cohort of Language Learners"
author: "Amy Anderson"
date: "2024-04-09"
output: html_document
---


#### Building the machinery for the social learning of multilingualism

The goal of this script is to generate a cohort of individuals who learn languages from each other in a multilingual space, in such a way that every individual develops a realistic language biography of whatever languages they are exposed to. This should form the basis of more complicated, age-structured, multi-generational models of the evolutionary dynamics of multilingualism. 
```{r, message=FALSE}
### Load libraries
library(tidyverse)
library(purrr)

# Load bespoke functions
source('./Functions/Pair agents for conversations.R')
source('./Functions/Inherit languages from parents.R')

```


 *** I bet there's a way to load all the files inside a folder using the source command. How though?



  
#### Vertical Inheritance 
Just to get things, started, have every individual in the cohort inherit a language from each of their parents. If both parents pass on the same language, then this individual begins life as a monolingual. If each parent passes on a different language, then this individual begins life as a bilingual. 
```{r}

# Set Cohort size
agents = 100

# Define the languages vector
languages <- c("A", "B", "C")

# Generate starting values by having each agent inherit one language from each parent (some agents will therefore start out monolingual; some will start out bilingual).
language_bank <- inherit_languages(agents, languages)

```

  
  
  

#### Set the rate of language acquisition as an age-dependent function:
```{r}
library(cowplot)
# generate a sample of 1,000 people with a uniform distribution of ages 0-70 years.
ages <- seq(from = 0, to = 70, by = 0.5)
ppl <- data.frame(age = sample(ages, 1000, replace = TRUE))


language_rate <- function(x, params){
  params$r0 * (1 - (1 / (1 + exp(-params$a * (x - params$tc - params$d))))) + 0.5
}



acquisition_rates <- data.frame(r0 = 4.5, # starting rate at age = 0
                                a = 0.5, # steepness of sigmoidal decay rate
                                d = 12, # x-axis position of sigmoidal curve's center
                                tc = 0) # starting age of language exposure


# calculate language acquisition rates for simulated people as a function of age
ppl$learning_rate <- sapply(ppl$age, language_rate, params = acquisition_rates)

# calculate predicted language proficiency at each age (cumulative sum of learning rate)
ppl <- ppl %>%
  arrange(age) %>%
  unique() %>%
  mutate(proficiency = cumsum(learning_rate))


rate_plot <- ggplot(ppl, aes(x = age, y = learning_rate)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 4.5)) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

prof_plot <- ggplot(ppl, aes(x = age, y = proficiency)) +
  geom_point(color = "blue") +
  geom_vline(xintercept = 10, linetype = 2) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()


plot_grid(rate_plot, prof_plot, ncol = 2)

```
  
   
   
#### To have agents learn from each other:  
 - Every tick of model time, agents are paired at random.
 - Each agent chooses (at random) to speak one of the languages in their language bank.
 - If their interlocutor speaks a language that is new to them, that language is added to their bank.
 - If they already know the language spoken by their interlocutor, then their proficiency in that language increases
 - The amount that language proficiency increases is moderated by the agent's age. 

  
  
The function below will **update the language bank** for every individual in the cohort based on their interaction with another agent. 
       
```{r}

### Define a FUNCTION to update everyone's language bank across multiple years of model time, as a function of age and social exposure to language. 

update_language_bank <- function(language_bank, pairing = "random", tmax, speaking = FALSE) {
  language_list <- list()
  for (t in seq(tmax)) { # for each new year in model time
    if (pairing == "random") {
      pair <- pair_agents_randomly(length(language_bank)) # pick conversational partners
    }
    for (pair_row in seq_len(nrow(pair))) {
      speaker1 <- pair[pair_row, 1]
      speaker2 <- pair[pair_row, 2]
      
      
      # Identify the languages currently known by each agent
      speaker1_languages <- names(purrr::keep(language_bank[[speaker1]], function(x) all(x > 0)))
      speaker2_languages <- names(purrr::keep(language_bank[[speaker2]], function(x) all(x > 0)))
  
      # Each speaker picks a language to speak —- sampled at random from each individual's language bank
      speaker1_says <- sample(speaker1_languages, 1)
      speaker2_says <- sample(speaker2_languages, 1)
      
      # The amount of proficiency gained is a function of the listener's age.
      params <- data.frame(d = 18, a = 0.5, r0 = 9, tc = 0) # parameter values chosen by simulating a monolingual cohort that achieves total proficiency in birth language by age 10. 
      age_factor <- params$r0 * (1 - (1 / (1 + exp(-params$a * (t - params$tc - params$d))))) + 0.5
      
      ### Learning by Listening ###
      # If speaker2 recognizes the language spoken by speaker1, then exposure to it improves their proficiency
      if (speaker1_says %in% speaker2_languages) {
        language_bank[[speaker2]][[speaker1_says]][t + 1] <- modify_depth(language_bank[[speaker2]][[speaker1_says]][t], 1, ~ .x + age_factor)
      } 
        else {
        # Add the new language as a named vector in speaker2's language bank
        language_bank[[speaker2]][[speaker1_says]][t + 1] <- 1
      }
      
      # If speaker1 recognizes the language spoken by speaker2, then exposure to it improves their proficiency
      if (speaker2_says %in% speaker1_languages) {
        language_bank[[speaker1]][[speaker2_says]][t + 1] <- modify_depth(language_bank[[speaker1]][[speaker2_says]][t], 1, ~ .x + age_factor)
      } 
       else {
        # Add the new language as a named vector in speaker1's language bank
        language_bank[[speaker1]][[speaker2_says]][t + 1] <- 1
      }
      
      if(speaking == FALSE){ #   If speaker1 does not recognize the language spoken by speaker2,
     # update all languages in speaker1's language bank (except the one spoken by speaker2)
     # to maintain their proficiency levels from time t to t+1
         for (lang in languages) {
    if (lang != speaker2_says) {
      language_bank[[speaker1]][[lang]][t + 1] <- language_bank[[speaker1]][[lang]][t]
    }
           if(lang != speaker1_says){
               language_bank[[speaker2]][[lang]][t + 1] <- language_bank[[speaker2]][[lang]][t]
               }
             }
           }

 ### Learning by Speaking ###
      if(speaking == TRUE){
        language_bank[[speaker1]][[speaker1_says]][t + 1] <- modify_depth(language_bank[[speaker1]][[speaker1_says]][t], 1, ~ .x + age_factor)
        language_bank[[speaker2]][[speaker2_says]][t + 1] <- modify_depth(language_bank[[speaker2]][[speaker2_says]][t], 1, ~ .x + age_factor)
        
     for (lang in languages) {
    if (lang != speaker2_says & lang != speaker1_says) {
      language_bank[[speaker1]][[lang]][t + 1] <- language_bank[[speaker1]][[lang]][t]
            language_bank[[speaker2]][[lang]][t + 1] <- language_bank[[speaker2]][[lang]][t]
           }
         }
     }
  }
  }
  language_bank <- modify_depth(.x = language_bank, .depth = -1, .f = \(x) if_else(x > 100, 100, x)) # impose a ceiling on proficiency values at proficiency = 100
    return(language_bank)
}



### Trial run of model function
starting_bank <- inherit_languages(agents, languages) # initial language bank, tracking all languages inherited vertically by infants in the cohort of agents
test <- update_language_bank(starting_bank, tmax = 70) # test run: 70 years of language learning


```

  


#### Design an output table for population-level results: What % of the cohort knows each language at time t?
The code below that creates a summary table was written by ChatGPT.  
It runs, but it's not particularly efficient. Something to optimize for future, if desired. 
```{r}

# Define a function to summarize language bank at each time point: e.g., Number of speakers of language A at time t
count_speakers <- function(language_bank, 
                           proficiency_threshold) {
  # Initialize an empty list to store summaries
  summary_list <- list()
  
  # Iterate over each time point
  for (t in seq_along(language_bank[[1]][[1]])) {
    # Initialize an empty data frame to store summary at current time point
    summary_df <- data.frame(Time = t)
    
    # Iterate over each language
    languages <- names(language_bank[[1]])
    for (lang in languages) {
      # Calculate count of agents with values > chosen proficiency threshold for current language at current time point
      count <- sum(map_lgl(language_bank, ~ .x[[lang]][t] > proficiency_threshold))
      
      # Add count to summary data frame
      summary_df[[lang]] <- count
    }
    
    # Append summary data frame to the summary list
    summary_list[[t]] <- summary_df
  }
  
  # Combine all summary data frames into a single data frame
  summary_table <- do.call(rbind, summary_list)
  
  return(summary_table)
}



```


Function to Plot number of speakers of each language over time: 
```{r}
library(wesanderson)

plot_speakers <- function(language_bank, proficiency){
  tables <- data.frame()
for(i in seq(proficiency)){
  summary_table <- count_speakers(language_bank, proficiency_threshold = proficiency[i]) # apply count_speakers function defined above
  summary_table$proficiency <- paste("proficiency >", proficiency[i], sep = " ")
  tables <- rbind(tables, summary_table)
}

speakers_plot <- pivot_longer(tables, cols = c(-"Time", -"proficiency"), names_to = "Language", values_to = "Speakers") %>%
  ggplot(aes(x = Time, y = Speakers)) +
  xlab("Age") +
  geom_line(aes(color = Language)) + 
  facet_wrap(~ proficiency) +
  scale_color_manual(values = wes_palette("Darjeeling1", type = "discrete")) +
  theme_minimal()
return(speakers_plot)
}

```


  Run some models with different parameters:
  
  
```{r}
proficiency <- c(0, 20, 50, 99)

p <- plot_speakers(test, proficiency)
save_plot('./Figures/cohort_models/speakercount_trilingual_random.png', p)

p

```
  
  
  In the monolingual cohort, 20% proficiency is reached at age 2. 
```{r}
just_listen <- update_language_bank(starting_bank, pairing = "random", tmax = 79, speaking = FALSE)

plot_speakers(just_listen, proficiency)
```

```{r}

listen_and_speak <- update_language_bank(starting_bank, pairing = "random", tmax = 79, speaking = TRUE)
plot_speakers(listen_and_speak, proficiency)

```
  Honestly, I'm not sure that 'learning by speaking' belongs in this model. Keep it to learning by exposure.
  
  

#### Plotting language biographies

```{r, echo=FALSE}

#### Reshape the language bank data for plotting
# Create a data frame to store the language bank information
language_bank <- test
df <- data.frame()
for(i in seq(language_bank)){
language_bio <- as_tibble(language_bank[[i]]) %>%
  mutate(agent_ID = i, .before = "A") %>%
  add_column(age = 1:nrow(.), .before = "A") %>%
  pivot_longer(cols = languages, names_to = "languages", values_to = "proficiency")
df <- rbind(df, language_bio)
}
df$proficiency <- unlist(df$proficiency)



plot_bios <- function(language_bank){
  df <- data.frame()
  languages <- names(language_bank[[1]])
for(i in seq(language_bank)){
language_bio <- as_tibble(language_bank[[i]]) %>%
  add_column(agent_ID = i, .before = "A") %>%
  add_column(age = 1:nrow(.), .before = "A") %>%
  pivot_longer(cols = languages, names_to = "languages", values_to = "proficiency")
df <- rbind(df, language_bio)
}
df$proficiency <- unlist(df$proficiency)
#### Plot the output!

p <- ggplot(df, aes(x = age, y = proficiency)) +
  geom_line(aes(color = as.factor(agent_ID))) +
  facet_wrap(~ languages) +
  theme_bw() +
  theme(legend.position = "none")
return(p)
}


p_bio <- plot_bios(test)
save_plot('./Figures/cohort_models/language_bios_trilingual_random.png', p_bio)
p_bio
```
  
  This plot shows language proficiency trajectories for each of these three hypothetical languages in a completely randomized environment. Each line is the lifetime language proficiency of a single agent in a single language. Agents may be exposed to 1, 2, or 3 languages over their lifetime.  
  

```{r}
bio_samples <- df %>%
  filter(agent_ID %in% 1:25) %>%
ggplot(aes(x = age, y = proficiency)) +
  geom_line(aes(color = languages)) +
  facet_wrap(~ agent_ID) +
  theme_bw() +
  theme(legend.position = "top")

save_plot('./Figures/cohort_models/sample_language_bios_trilingual_random.png', bio_samples)
bio_samples

```
  
  This plot shows a sample of individual language biographies for agents in the cohort. Right now the rate of language learning and its relationship to individual age are not calibrated using any empirical data; it's a fairly arbitrary choice, just to focus on getting the guts of the code to run.  










```{r}

```

