---
title: "Model 2.0 - Multi-generational Multilingualism"
author: "Amy Anderson"
date: "2024-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
### Load libraries
library(tidyverse)
library(cowplot)

# Load bespoke functions
function_files <- list.files(path="./Functions/", full.names=TRUE)
sapply(function_files, FUN=source) # this command tells me that 'object 'generate_agent_id' not found', even though calling that function file directly on its own works just fine. Not sure how to address this error. 
source('./Functions/Generate Agent IDs.R')
```



Generate a starting set of agents, and their age structure:
```{r}
agent_census <- data.frame(agent_id = sapply(seq(from = 0, length.out = 1000), FUN = generate_agent_id))

# uniform age structure
agent_census$age <- sample(0:80, 1000, replace = TRUE)

# alternate assigning male and female state for each agent. 
agent_census$female <- rep(c(0,1), nrow(agent_census)/2)
```


Pair up males/females for reproductive partnerships: 
```{r}

```


Assign a family number to each couple.
...This is going to get messed up after one generation....
No problem!
Make each agent's Family ID status a vector that can contain multiple family IDs 
Family IDs will generate for each new marriage-partnered couple. 
All agents will need to get married at age 15 then...
```{r}

```




