---
title: "Model 2.0 - Multi-generational Multilingualism"
author: "Amy Anderson"
date: "2024-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
### Load libraries
library(tidyverse)
library(cowplot)



### Load bespoke functions

## create character vector of all file names inside the 'Functions' folder
# specify function to generate agent IDs first -- this function is called in the 'Example Use' section of some other function files, and that will cause issues if it isn't loaded into the workspace before the files that use it.
first_function <- "./Functions//Generate Agent IDs.R"
function_files <- list.files(path="./Functions/", full.names=TRUE) # all file names in Function folder
function_files <- function_files[which(function_files != "./Functions//Generate Agent IDs.R")] # remove the agent ID function
function_files <- c(first_function, function_files) # put it back in as the first item in the list of names. 

# apply the source() command to all Function file names them to pull the functions into the working environment. 
sapply(function_files, FUN=source) 

```



Generate a starting set of agents, and their age structure:
```{r}
agent_census <- data.frame(agent_id = sapply(seq(from = 0, length.out = 1000), FUN = generate_agent_id))

# uniform age structure
agent_census$age <- sample(0:80, 1000, replace = TRUE)

# alternate assigning male and female state for each agent. 
agent_census$female <- rep(c(0,1), nrow(agent_census)/2)
```


Pair up males/females for reproductive partnerships: 
```{r}

```


Assign a family number to each couple.
...This is going to get messed up after one generation....
No problem!
Make each agent's Family ID status a vector that can contain multiple family IDs 
Family IDs will generate for each new marriage-partnered couple. 
All agents will need to get married at age 15 then...
```{r}

```


# Designate languages in play. 
```{r}
choose_languages <- function(number_of_languages){
  languages <- chartr("123456789", "ABCDEFGHI", seq(number_of_languages))
return(languages)
}

languages <- choose_languages(3)
```

