---
title: "Fine tuning the rate of language learning"
author: "Amy Anderson"
date: "2024-05-01"
output: html_document
---

 Assume that an agent who inherits a language from their parent and is exposed to the same language every round until age 10 reaches 100% proficiency in that language.
 

```{r, message=FALSE}

### Load libraries
library(tidyverse)
library(purrr)
library(cowplot)



# Load bespoke functions
source('./Functions/Pair agents for conversations.R')
source('./Functions/Inherit languages from parents.R')

```
 
 
 
```{r}

# Set Cohort size
agents = 100

# Define the languages vector
languages <- c("A")

# Generate starting language proficiency values for monolingual cohort
mono_babies <- inherit_languages(agents, languages)

```
 
 
 
```{r}
#### Set the rate of language acquisition as an age-dependent function:

language_rate <- function(x, params){
  params$r0 * (1 - (1 / (1 + exp(-params$a * (x - params$tc - params$d))))) + 0.5
}

```



### Parameters for language acquisiton rate function ###
```{r}

# starting rate at age = 0
r0 = 5
r0_sweep <- c(1, 3, 5, 9)
# steepness of sigmoidal decay rate
a = c(0.5)
a_sweep <- c(0.25, 0.5, 0.75, 1)
# x-axis position of sigmoidal curve's center
d = 12
d_sweep <- c(8, 12, 14, 18)
# starting age of language exposure
tc = 0
tc_sweep <- c(0, 2, 4, 8)

acquisition_rates <- data.frame(r0_sweep, a, d, tc) 


# calculate language acquisition rates at each age
ages <- seq(from = 0, to = 50, by = 1)

learning_rates <- tibble(ages)
for(i in 1:nrow(acquisition_rates)){
  var[i] <- paste("learning_rate_r0", acquisition_rates$r0[i], sep = "")
  rates <- sapply(ages, language_rate, params = acquisition_rates[i,])
  learning_rates <- cbind(learning_rates, rates)
}
names(learning_rates) <- c("ages", var)

df <- learning_rates %>%
  pivot_longer(cols = -"ages", names_to = "r0", values_to = "learning_rate") %>%
  group_by(r0) %>%
  mutate(proficiency = cumsum(learning_rate)) %>% # calculated accumulated language proficiency by each age
  ungroup()




rate_plot <- ggplot(df, aes(x = ages, y = learning_rate)) +
  geom_line(aes(color = r0)) +
 # facet_wrap(~ r0, ncol = length(unique(df$r0))) +
#  scale_y_continuous(limits = c(0, 4.5)) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

prof_plot <- ggplot(df, aes(x = ages, y = proficiency)) +
  geom_line(aes(color = r0)) +
 # facet_wrap(~ r0, ncol = length(unique(df$r0))) +
  geom_vline(xintercept = 10, linetype = 2) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

plot_grid(rate_plot, prof_plot, nrow = 2)

```
  
  This plot shows the effect of changing the starting value for age-specific rate of language learning on age-specific values for language proficiency. 
  
  
  
  Test the sensitivity of proficiency to changing the value for a, the steepness of the sigmoidal decay rate:  
```{r}

acquisition_rates <- data.frame(r0, a_sweep, d, tc) 

learning_rates <- tibble(ages)
for(i in 1:nrow(acquisition_rates)){
  var[i] <- paste("learning_rate_a", acquisition_rates$a[i], sep = "")
  rates <- sapply(ages, language_rate, params = acquisition_rates[i,])
  learning_rates <- cbind(learning_rates, rates)
}
names(learning_rates) <- c("ages", var)

df <- learning_rates %>%
  pivot_longer(cols = -"ages", names_to = "a", values_to = "learning_rate") %>%
  group_by(a) %>%
  mutate(proficiency = cumsum(learning_rate)) %>% # calculated accumulated language proficiency by each age
  ungroup()

prof_at_10 <- df %>% 
  filter(ages == 10)


rate_plot <- ggplot(df, aes(x = ages, y = learning_rate)) +
  geom_line(aes(color = a)) +
#  facet_wrap(~ a, ncol = length(unique(df$a))) +
#  scale_y_continuous(limits = c(0, 4.5)) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

prof_plot <- ggplot(df, aes(x = ages, y = proficiency)) +
  geom_line(aes(color = a)) +
#  facet_wrap(~ a, ncol = length(unique(df$a))) +
  geom_vline(xintercept = 10, linetype = 2) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

plot_grid(rate_plot, prof_plot, nrow = 2)
```
  
  
  Test the effect of changing d, the x-axis position of the sigmoidal curve's center

```{r}

acquisition_rates <- data.frame(r0, a, d_sweep, tc) 

learning_rates <- tibble(ages)
for(i in 1:nrow(acquisition_rates)){
  var[i] <- paste("learning_rate_d", acquisition_rates$d[i], sep = "")
  rates <- sapply(ages, language_rate, params = acquisition_rates[i,])
  learning_rates <- cbind(learning_rates, rates)
}
names(learning_rates) <- c("ages", var)

df <- learning_rates %>%
  pivot_longer(cols = -"ages", names_to = "d", values_to = "learning_rate") %>%
  group_by(d) %>%
  mutate(proficiency = cumsum(learning_rate)) %>% # calculated accumulated language proficiency by each age
  ungroup()

prof_at_10 <- df %>% 
  filter(ages == 10)


rate_plot <- ggplot(df, aes(x = ages, y = learning_rate)) +
  geom_line(aes(color = d)) +
#  facet_wrap(~ a, ncol = length(unique(df$a))) +
#  scale_y_continuous(limits = c(0, 4.5)) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

prof_plot <- ggplot(df, aes(x = ages, y = proficiency)) +
  geom_line(aes(color = d)) +
#  facet_wrap(~ a, ncol = length(unique(df$a))) +
  geom_vline(xintercept = 10, linetype = 2) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

plot_grid(rate_plot, prof_plot, nrow = 2)
```

 
 
 
```{r}

```
 
```{r}
d = 18
a = 0.5
r0 = 9
tc = 0

params <- data.frame(d, a, r0, tc)
learning_rate <- sapply(ages, language_rate, params = params)
df <- data.frame(ages, learning_rate) %>%
  mutate(proficiency = cumsum(learning_rate))


rate_plot <- ggplot(df, aes(x = ages, y = learning_rate)) +
  geom_line() +
#  facet_wrap(~ a, ncol = length(unique(df$a))) +
#  scale_y_continuous(limits = c(0, 4.5)) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

prof_plot <- ggplot(df, aes(x = ages, y = proficiency)) +
  geom_line(color = "blue") +
#  facet_wrap(~ a, ncol = length(unique(df$a))) +
  geom_vline(xintercept = 10, linetype = 2) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

plot_grid(rate_plot, prof_plot, nrow = 2)
```
 
 
 
Check out the proficiency curve for language immersion starting at different ages:

```{r}

acquisition_rates <- data.frame(r0, a, d, tc_sweep) 

learning_rates <- tibble(ages)
for(i in 1:nrow(acquisition_rates)){
  var[i] <- paste("age_at_immersion_", acquisition_rates$tc[i], sep = "")
  rates <- sapply(ages, language_rate, params = acquisition_rates[i,])
  learning_rates <- cbind(learning_rates, rates)
}
names(learning_rates) <- c("ages", var)

df <- learning_rates %>%
  pivot_longer(cols = -"ages", names_to = "tc", values_to = "learning_rate") %>%
  group_by(tc) %>%
  mutate(proficiency = cumsum(learning_rate)) %>% # calculated accumulated language proficiency by each age
  ungroup()

prof_at_10 <- df %>% 
  filter(ages == 10)


rate_plot <- ggplot(df, aes(x = ages, y = learning_rate)) +
  geom_line(aes(color = tc)) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

prof_plot <- ggplot(df, aes(x = ages, y = proficiency)) +
  geom_line(aes(color = tc)) +
  geom_vline(xintercept = 10, linetype = 2) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_bw()

plot_grid(rate_plot, prof_plot, nrow = 2)
```

... I don't think the tc term in the equation is doing what it says it's doing. Not sure it's useful, given the bottom-up approach of this project. 



test run monolingual language acquisition
```{r}
testrun <- update_language_bank(mono_babies, tmax = 50)



plot_bios <- function(language_bank){
  df <- data.frame()
  languages <- names(language_bank[[1]])
for(i in seq(language_bank)){
language_bio <- as_tibble(language_bank[[i]]) %>%
  add_column(agent_ID = i, .before = "A") %>%
  add_column(age = 1:nrow(.), .before = "A") %>%
  pivot_longer(cols = languages, names_to = "languages", values_to = "proficiency")
df <- rbind(df, language_bio)
}
df$proficiency <- unlist(df$proficiency)
#### Plot the output!

p <- ggplot(df, aes(x = age, y = proficiency)) +
  geom_line(aes(color = as.factor(agent_ID))) +
  facet_wrap(~ languages) +
  theme_bw() +
  theme(legend.position = "none")
return(p)
}


plot_bios(testrun)
```

 Ok, there we have it -- these model parameters produce agents who reach 100% monolingual language proficiency at age 12. 